### 2019-11-01 paper/paper.tex
Adds Zotero-generated bibfile

Spatial patterns in neural connectivity provide clues about the constraints
under which brains evolve and develop. Whether or not patterning governs
information processing, models of the development of information processing
circuits can be validated in terms of their ability to account for biological
patterning (\citealp{Purves1992}; see also
\citealp{Wilson2015,Bednar2016}). One of the most distinctive patterns of
neural connectivity is found in the rodent barrel cortex
(\citealp{Woolsey1948,Woolsey1970,Welker1974}, see
\citealp{Fox2008}). Understanding how this particular pattern forms might shed
light on the interplay of genetic and extrinsic factors that shape cortical
circuits, and on a wealth of anatomical, physiological, and behavioural data
obtained using this model system.

@@

Barrel formation requires afferent input, driven by whisker stimulation and by
thalamic calcium waves that propagate via mGlu5-receptor activation
(\citealp{Anton2019}, see also \citealp{Anton2018}). Patterning also depends
on a complex network of axon guidance processes, which includes interactions
between guidance molecules like ephrin-A5 and A7 \citep{Miller2006} and
between adhesion molecules like cadherin-6 and 8 \citep{Greig2013,Bishop2000},
and the binding of p75 to neurotropins (\citealp{Shimogori2005,Bishop2002};
see also \citealp{Dye2011a,Dye2011b}). This network is orchestrated by
interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2,
Pax6, Sp8, and Coup-tf1 \citep{Shimogori2005,Bishop2000}, which are expressed
in gradients that mark orthogonal axes and can be manipulated to cause the
barrels to stretch, shrink, shift, skew, and even duplicate
\citep{Shimogori2005,Assimacopoulos2012,Borello2014,Sahara2007,Ypsilanti2016,Sur2005}.

@@

\begin{equation} \label{eq:Karb2DExt_dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{i'=1}^{N} c_{i'}(\mb{x}, t)\right)[a_i(\mb{x},t)]^k.
\end{equation}

@@

%
\begin{equation} \label{eq:Karb2DExt_da}
\frac{\partial a_i(\mb{x},t)}{\partial t} = \nabla\vcdot\left(D \nabla a_i(\mb{x},t)-a_i(\mb{x},t)\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i(\mb{x},t)}{\partial t} + \chi.
\end{equation}

@@

% Include figure ../boundaries/barreloids_vanderloos_graph.png
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{../boundaries/barreloids_vanderloos_graph.png}
\end{center}
\vspace{-2em}
\caption{Barreloid centres taken from \cite{van_der_loos_1976}, Fig.~2
  are shown as black circles. It is assumed that there are at least
  two linearly increasing, orthogonal signalling gradients in the
  thalamus (side graphs - addme!), such that the location of each
  barreloid within the graded thalamic signal field can differentiate
  the efferent axons, with the result that axons from each barreloid
  interact to varying degrees with cortical signalling
  gradients. Coloured arrows show, for each barreloid, the magnitude
  of the interaction that its thalamocortical projection will have
  with the four signalling gradients in the cortex. The mappings from
  barreloid location to interaction strength are linear.}
\label{fig:barreloids}
\end{figure*}

@@

We first verified that the system given by Eqns. \ref{eq:Karb2DExt_dc} \&
\ref{eq:Karb2DExt_da} described a two dimensional extension of the axon
branching model in \citet{Karbowski2004}. We showed that in the presence of
three guidance gradients varying only in one dimension (along the long axis of
an elliptical domain), five thalamocortical axon types, each interacting
uniquely with the three guidance gradients, would segregate into five separate
regions. We investigated a number of configurations of the guidance gradients,
varying their shape, $\rho(\mb{x})$, and number, $M$, and their effect on a
varying number of thalamocortical projections, $N$. In general, we were not
able to produce structure in the resulting patterns of the connections,
$c(\mb{x},t\rightarrow\infty)$, which was not present in the guidance
patterns.

Complex structure can indeed be observed in protein expression for some genes,
such as the molecules Bhlhb5 \citep{Joshi2008} and Lmo4
\citep{Cederquist2013,Greig2013}. These expression patterns can therefore
provide a theoretical mechanism for the guidance of axons into distinct
regions, shifting the explanation of how such structures are generated into
the realm of gene transcriptional interactions and mRNA diffusion/movement
(\cmnt{Fixme: find a ref on structure in gene transcription}). However,
complex, stable structure in gene expression is less commonly reported than
homogeneous or simply varying expression, such as Pax6, Emx2, Sp8 and Coup-tf1 which
display graded expression across the entire isocortex
(\cmnt{REFS}). Furthermore, the complex structure observed in the whisker
barrel field has not, to our knowledge, been linked to a complex set of gene
expression patterns matching the scale of the barrels. To investigate how the
whisker barrel structure might form in the presence of only two
coordinate-giving morphogen gradients, we turned our attention to the
competition for connections built into the model.

There exists an element of competition in the model system specified by
Eqns. \ref{eq:Karb2DExt_dc} \& \ref{eq:Karb2DExt_da}. The term $\left(1 -
\sum_{i=1}^{N} c_i(\mb{x}, t)\right)$ ensures that once the available synaptic
receptors in a region have been fully occupied, say by thalamocortical types
$i=2$ and $i=3$, there is no longer the possibility for type $i=1$ to form any
synapses. This competition turns out to be relatively weak, because it does
not strongly affect the branching density in that region; two or more TC axon
types can be branching equally in the region, leading to roughly equal numbers
of connections being made up until the connection limit is reached.

@@

Formation of a grid of domains, with somatotopic ordering.

Fig of domains.

@@

\bibliography{bibfile}

@@

%
\begin{equation} \label{eq:Karb2DExt_dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:Karb2DExt_conserve}
n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
\end{equation}
%
\begin{equation} \label{eq:Karb2DExt_da}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}- \frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
\end{equation}

>>>>

Spatial patterns in neural connectivity provide clues about the constraints
under which brains evolve and develop. Whether or not patterning governs
information processing, models of the development of information processing
circuits can be validated in terms of their ability to account for biological
patterning (\citealp{purves_iterated_1992}; see also
\citealp{wilson_what_2015,bednar_cortical_2016}). One of the most distinctive patterns of
neural connectivity is found in the rodent barrel cortex
(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974}, see
\citealp{Fox2008}). Understanding how this particular pattern forms might shed
light on the interplay of genetic and extrinsic factors that shape cortical
circuits, and on a wealth of anatomical, physiological, and behavioural data
obtained using this model system.

@@

Barrel formation requires afferent input, driven by whisker stimulation and by
thalamic calcium waves that propagate via mGlu5-receptor activation
(\citealp{Anton2019}, see also \citealp{Anton2018}). Patterning also depends
on a complex network of axon guidance processes, which includes interactions
between guidance molecules like ephrin-A5 and A7 \citep{Miller2006} and
between adhesion molecules like cadherin-6 and 8 \citep{greig_molecular_2013,Bishop2000},
and the binding of p75 to neurotropins (\citealp{Shimogori2005,Bishop2002};
see also \citealp{Dye2011a,Dye2011b}). This network is orchestrated by
interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2,
Pax6, Sp8, and Coup-tf1 \citep{Shimogori2005,Bishop2000}, which are expressed
in gradients that mark orthogonal axes and can be manipulated to cause the
barrels to stretch, shrink, shift, skew, and even duplicate
\citep{Shimogori2005,Assimacopoulos2012,Borello2014,Sahara2007,Ypsilanti2016,Sur2005}.

@@

\begin{equation} \label{eq:dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{i'=1}^{N} c_{i'}(\mb{x}, t)\right)[a_i(\mb{x},t)]^k.
\end{equation}

@@

%
\begin{equation} \label{eq:da}
\frac{\partial a_i(\mb{x},t)}{\partial t} = \nabla\vcdot\left(D \nabla a_i(\mb{x},t)-a_i(\mb{x},t)\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i(\mb{x},t)}{\partial t} + \chi.
\end{equation}

@@

@@

We first verified that the system given by Eqns. \ref{eq:dc} \&
\ref{eq:da} described a two dimensional extension of the axon
branching model in \citet{Karbowski2004}. We showed that in the presence of
three guidance gradients varying only in one dimension (along the long axis of
an elliptical domain), five thalamocortical axon types, each interacting
uniquely with the three guidance gradients, would segregate into five separate
regions. We investigated a number of configurations of the guidance gradients,
varying their shape, $\rho(\mb{x})$, and number, $M$, and their effect on a
varying number of thalamocortical projections, $N$. In general, we were not
able to produce structure in the resulting patterns of the connections,
$c(\mb{x},t\rightarrow\infty)$, which was not present in the guidance
patterns.

Complex structure can indeed be observed in protein expression for some genes,
such as the molecules Bhlhb5 \citep{Joshi2008} and Lmo4
\citep{Cederquist2013,greig_molecular_2013}. These expression patterns can therefore
provide a theoretical mechanism for the guidance of axons into distinct
regions. The question of how such structures are generated becomes one of of
how gene transcriptional interactions and mRNA diffusion and movement achieves
this (\cmnt{Fixme: find a ref on structure in gene transcription}). However,
complex and stable structure in gene expression data is commonly reported;
monotonically varying expression patterns, such as those for Pax6, Emx2, Sp8
and Coup-tf1 which display graded expression across the entire isocortex
(\cmnt{REFS}) are more typical. Furthermore, the complex structure observed in
the whisker barrel field has not, to our knowledge, been linked to a complex
set of gene expression patterns matching the scale of the barrels. To
investigate how the whisker barrel structure might form in the presence of
only monotonically varying morphogen gradients, we turned our attention to the
competition for connections built into the model.

There exists an element of competition in the model system specified by
Eqns. \ref{eq:dc} \& \ref{eq:da}. The term $\left(1 -
\sum_{i=1}^{N} c_i(\mb{x}, t)\right)$ in Eq.~\ref{eq:dc} ensures that once the available synaptic
receptors in a region have been fully occupied, say by thalamocortical types
$i=2$ and $i=3$, there is no longer the possibility for type $i=1$ to form any
synapses. This competition turns out to be relatively weak, because it does
not strongly affect the branching density in that region; two or more TC axon
types can be branching equally in the region, leading to roughly equal numbers
of connections being made up until the connection limit is reached.

@@

% Include figure ../boundaries/barreloids_vanderloos_graph.png
\begin{figure*}
\begin{center}
\includegraphics[width=0.7\textwidth]{../boundaries/barreloids_vanderloos_graph.png}
\end{center}
%\vspace{-0em}
\caption{Barreloid centres taken from \cite{van_der_loos_1976}, Fig.~2
  are shown as black circles. It is assumed that there are at least
  two linearly increasing, orthogonal signalling gradients in the
  thalamus (side graphs - addme!), such that the location of each
  barreloid within the graded thalamic signal field can differentiate
  the efferent axons, with the result that axons from each barreloid
  interact to varying degrees with cortical signalling
  gradients. Coloured arrows show, for each barreloid, the magnitude
  of the interaction that its thalamocortical projection will have
  with the four signalling gradients in the cortex. The mappings from
  barreloid location to interaction strength are linear.}
\label{fig:barreloids}
\end{figure*}

@@

\bibliography{BarrelEmerge}

@@

%
\begin{equation} \label{eq:dc2}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:conserve}
n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
\end{equation}
%
\begin{equation} \label{eq:da2}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}- \frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
\end{equation}
